#Individual Assignment for Statistical Learning Class
#AUTHOR:ANIL SENER
#The purpose of this assignment was to use clustering methodsin telecommunication radio #network daily timeseries data (minute of day versus connection period of each mobile phone to the base station) from 50 different base stations to identify whether the base #station is located in near a road/ railway or not. I am keeping the data anonymous to #secure the data protection rights of my provider. However this nice piece of R code #achieved %100 success with a simple metholodology exploiting descriptive stats. (Although #I was thinking I missed one base station) Another methodology might be using Time Series #Clustering with advanced distance measures but it would be very complicated and slower #algorithm to handle tens of thousands of documents daily.
#Potential use cases that I can think of are smart city planning and administration,
#scheduling in transportation industry,advertising optimization etc. 
#The results are provided in the image files


#Storing files to different dataframes
```{r}
require(plyr)
require(dostats)
library(TTR)
library(psych)
files <- list.files(pattern = ".csv", full.names=FALSE)

df_list<-list()
for (i in seq_along(files)) {

    assign(paste(files[[i]]), read.csv(files[i],sep=";",header=FALSE))

    df_list[[i]]<-assign(paste(paste(files[[i]], i, sep = ""), "summary", sep = "."), 
           ldply(get(paste(files[[i]])), dostats, sum, min, mean, median, max,sd,skew,mad))
 
}

```

#Building new dataframe based on descriptive stats summary of no of minutes to jump from one antenna to another for each file

```{r}
df <- data.frame()
for (i in seq_along(df_list)) {
  df <- rbind(df, df_list[[i]][df_list[[i]]$.id=="V2",])
  
}
df<-cbind(df,files)
```



#Plotting Standard deviations of each file 
```{r}
library(ggplot2)

ggplot (aes(x=files, y=sd), data=df) +
    geom_point()
```
#compared to other features sd seems to provide a better distinction between files


#Executing k-means clustering algorithm
```{r}
km.out <- kmeans(df$sd,2,nstart=20)

# append cluster assignment
df <- data.frame(df, km.out$cluster)
km.out$withinss
km.out$betweenss/km.out$totss

```
#Within cluster sum of squares by cluster:
#[1] 52262.88 51767.15
#between_SS / total_SS =  [1] 0.8128258 %


#plotting the SD instances for k-means clustering 
```{r}
ggplot (aes(x=files, y=sd), data=df) +
    geom_point(color=df$km.out.cluster)

```

#Probing hcluster
```{r}
hc.complete <- hclust(dist(df$sd), method="complete")
plot(hc.complete)
hcluster <- cutree(hc.complete, 2)
df <- data.frame(df, hcluster)
```


#plotting the SD instances for hclustering 
```{r}
ggplot (aes(x=files, y=sd), data=df) +
    geom_point(color=df$hcluster)

```




#Assigning clusters to the full dataset
```{r}
filenames <- list.files(pattern="*.csv", full.names=FALSE)

for (i in seq_along(filenames)){
       
  # if the merged dataset doesn't exist, create it
  if (!exists("dataset")){
    dataset <- read.table(filenames[[i]], header=FALSE, sep=";")
    dataset <- cbind(dataset,filenames[[i]])
    dataset <- cbind(dataset,km.out$cluster[[i]])
    dataset <- cbind(dataset,hcluster[[i]])
  }
   
  # if the merged dataset does exist, append to it
  if (exists("dataset")){
    temp_dataset <-read.table(filenames[[i]], header=FALSE, sep=";")
    temp_dataset <- cbind(temp_dataset,filenames[[i]])
    temp_dataset <- cbind(temp_dataset,km.out$cluster[[i]])
    temp_dataset <- cbind(temp_dataset,hcluster[[i]])
    dataset<-rbind(dataset, temp_dataset)
    rm(temp_dataset)
  }
 
}
colnames(dataset)[3] <- "filename"
colnames(dataset)[4] <- "kcluster"
colnames(dataset)[5] <- "hcluster"
```

#Plotting each file based on cluster it is belonging to (kmeans)
```{r}
ggplot (aes(x=V1, y=V2, color=factor(kcluster)), data=dataset) +
    geom_point(size=1)+ facet_wrap(~filename)
```
#Plotting each file based on cluster it is belonging to (hcluster)

```{r}
ggplot (aes(x=V1, y=V2, color=factor(hcluster)), data=dataset) +
    geom_point(size=1)+ facet_wrap(~filename)
```


#Exporting the road & random labels to an excel file
```{r}
output<-as.data.frame(cbind(as.vector(filenames),as.vector(hcluster)))
output$label[output$V2 == 1]<-"random"
output$label[output$V2 == 2]<-"road"
output<-output[,2:3]
colnames(output)[1]<-"file"
colnames(output)[2]<-"label"
library(xlsx)
write.xlsx(output, "output.xlsx")
```

Summary:
Hierarchical Clustering with euclidian distance provided the best result in labelling the roads (24 out of 25). I have also tried all other distance methods, most of them provided similar results. I also wanted to use mahalanobis distance however I couldn't make mahalanobis() function to work for singular vector with and inverted covariance matrix. Labelling scheme is organized as "random" or "road" format. 
